---
title: "WUP_04"
author: "Chi N"
date: "summer2024"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata<-read.csv(file="GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
basicCN <- ggplot(mydata,aes(x=DEGREE,y=WORDSUM))
```
```{r}
basicCN + geom_boxplot()
```

By looking at this data the graduate and junior college boxes were above high school as well as Lt high school. The graduate and bachelor boxed would bump into each other that they are close. Graduate school and Lt high school are different. Graduate and high school are different as well. 

### Numerical Descriptive Results
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```

When looking at the data, i see a huge difference in student in high school to junior college. High school students were a total of 85 whereas junior college had the least amount at 12. The students with a bachelor degree were the second highest at 44 where the students in LT hs were at 16. All of the students attending had either a lot or a little bit of students. A big disparity between the number of students. There is going to be some statistical difference between these groups, it is not completely clear to what is different statistically. To make it clear, we will have to run these statistical tests. There will be a two part test, first needing to run one way anova, to show us some statistical difference. To clarify this we would need to run the turkey HSD.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```

The probability of getting these results is significant in the 1/1000s level. This means we need to find out the differences. 


```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 15.7, 2, 2))
plot(TukeyHSD(model1,conf.level=.98),las=2)
```

What is difference is high school and LT school. They are both different from graduate school and nothing else is different. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

We struggled but finally figured out how to change it from .95 which is default to .98 which showed that high school and  junior college were not different at 98% level. Even if i were to go to the .95 percent level you would find that they were different. 
